---
title: "Analyses for Aim 1 of \"Pregnancy during the pandemic: The impact of COVID-19-related stress on risk for prenatal depression\""

author: "Lucy S. King"
date: "2020"
output: 
  github_document:
    toc: true
    toc_depth: 2
---

# Set up environment
```{r}
# Libraries
library(tidyverse)
library(lubridate)
library(parameters)
library(ggsci)
library(ggpubr)
library(relaimpo)
library(effectsize)
library(MatchIt)
library(cem)
library(codebook)
library(labelled)
source("R_rainclouds.R")

# Data files

## BABIES
babies_data_home <- "~/Box/Mooddata_Coordinating/BABIES/Data/final_scored_data/"
babes_demo_file <- paste0(babies_data_home, "demographics/demo_longitudinal_cleaned_final.csv")
babies_epds_file <- paste0(babies_data_home, "EPDS/epds_wf_complete.csv")
babies_prams_file <- paste0(babies_data_home, "PRAMS/prams_alldata_cleaned.csv")
babies_treatment_file <- paste0(babies_data_home, "SCID/scid_lg_preg_treatment_history_complete.csv")

# ## COPE
cope_data <- "~/Box/Mooddata_Coordinating/COPE/data/wrangled_cleaned_data_PREG_20201118.csv"

set.seed(100588)

```

# Read in data

## BABIES data
```{r}
babies <-
  read_csv(babes_demo_file) %>% 
  left_join(
    read_csv(babies_treatment_file), by = "ID"
  ) %>% 
  dplyr::select(
    ID,
    mom_age_t1,
    education = education_t1_txt,
    income_bins = annual_income_t1_txt,
    income_numeric = annual_income_t1_num,
    income_needs_t1, 
    employment_status_t1_txt,
    employment_explain_t1,
    pregnant_times_t1,
    pregnant_times_t2,
    pregnant_times_t3,
    current_num_child_t1 = current_num_children_t1,
    gestational_weeks_t1,
    ppl_in_home_adults_t1,
    ppl_in_home_allchild_t1,
    ppl_in_home_tot_t1,
    mom_race,
    mom_latinx,
    marital_status_t1_txt,
    past_mh_sa_treatment
  ) %>% 
  mutate(
    pregnant_times = if_else(
      is.na(pregnant_times_t1), pregnant_times_t2, pregnant_times_t1
    ),
    pregnant_times = if_else(
      is.na(pregnant_times), pregnant_times_t3, pregnant_times
    )
  ) %>% 
  dplyr::select(
    -pregnant_times_t1:-pregnant_times_t3
  ) %>% 
  rename_at(
    vars(-ID),
    funs(str_replace(., "_t1", ""))
  ) %>% 
  left_join(
    read_csv(babies_epds_file) %>% 
      dplyr::select(-epds_t3, -epds_t3_concern),
    by = "ID"
  ) %>% 
  mutate(cohort = "Pre-pandemic") %>% 
  dplyr::select(
    ID,
    cohort,
    epds_total = epds_t1,
    everything()
  ) %>% 
  mutate(
    job_status = case_when(
      employment_status_txt == "Out of work, looking for work" ~ "Unemployed",
      employment_status_txt == "Out of work, not looking for work" ~ "Unemployed",
      TRUE ~ employment_status_txt
    ),
    ID = paste0(ID, "_b"),
    first_pregnancy = if_else(
      !is.na(pregnant_times) & pregnant_times == 1, 1, 0
    ),
    current_num_child = if_else(
      is.na(current_num_child), ppl_in_home_allchild, current_num_child
    )
  ) %>% 
  #1078 did not actually participate in pregnancy; need to remove from BABIES pregnancy demographics
  filter(!is.na(epds_total))
```

```{r}
babies <-
  babies %>% 
  group_by(ID) %>% 
   # inr based on the federal poverty thresholds for number of people in household and number that are children
  mutate(
    fpl_threshold = case_when(
      ppl_in_home_tot == 1 & ppl_in_home_allchild == 0 ~ 13300,
      ppl_in_home_tot == 2 & ppl_in_home_allchild == 0 ~ 17120,
      ppl_in_home_tot == 2 & ppl_in_home_allchild == 1 ~ 17622,
      ppl_in_home_tot == 3 & ppl_in_home_allchild == 0 ~ 19998,
      ppl_in_home_tot == 3 & ppl_in_home_allchild == 1 ~ 20578,
      ppl_in_home_tot == 3 & ppl_in_home_allchild == 2 ~ 20598,
      ppl_in_home_tot == 4 & ppl_in_home_allchild == 0 ~ 26370,
      ppl_in_home_tot == 4 & ppl_in_home_allchild == 1 ~ 26801,
      ppl_in_home_tot == 4 & ppl_in_home_allchild == 2 ~ 25926,
      ppl_in_home_tot == 4 & ppl_in_home_allchild == 3 ~ 26017,
      ppl_in_home_tot == 5 & ppl_in_home_allchild == 0 ~ 31800,
      ppl_in_home_tot == 5 & ppl_in_home_allchild == 1 ~ 32263,
      ppl_in_home_tot == 5 & ppl_in_home_allchild == 2 ~ 31275,
      ppl_in_home_tot == 5 & ppl_in_home_allchild == 3 ~ 30510,
      ppl_in_home_tot == 5 & ppl_in_home_allchild == 4 ~ 30044,
      ppl_in_home_tot == 6 & ppl_in_home_allchild == 0 ~ 36576,
      ppl_in_home_tot == 6 & ppl_in_home_allchild == 1 ~ 36721,
      ppl_in_home_tot == 6 & ppl_in_home_allchild == 2 ~ 35965,
      ppl_in_home_tot == 6 & ppl_in_home_allchild == 3 ~ 35239,
      ppl_in_home_tot == 6 & ppl_in_home_allchild == 4 ~ 34161,
      ppl_in_home_tot == 6 & ppl_in_home_allchild == 5 ~ 33522,
      ppl_in_home_tot == 7 & ppl_in_home_allchild == 0 ~ 42085,
      ppl_in_home_tot == 7 & ppl_in_home_allchild == 1 ~ 42348,
      ppl_in_home_tot == 7 & ppl_in_home_allchild == 2 ~ 41442,
      ppl_in_home_tot == 7 & ppl_in_home_allchild == 3 ~ 40811,
      ppl_in_home_tot == 7 & ppl_in_home_allchild == 4 ~ 39635,
      ppl_in_home_tot == 7 & ppl_in_home_allchild == 5 ~ 38262,
      ppl_in_home_tot == 7 & ppl_in_home_allchild == 6 ~ 36757,
      ppl_in_home_tot == 8 & ppl_in_home_allchild == 0 ~ 47069,
      ppl_in_home_tot == 8 & ppl_in_home_allchild == 1 ~ 47485,
      ppl_in_home_tot == 8 & ppl_in_home_allchild == 2 ~ 46630,
      ppl_in_home_tot == 8 & ppl_in_home_allchild == 3 ~ 45881,
      ppl_in_home_tot == 8 & ppl_in_home_allchild == 4 ~ 44818,
      ppl_in_home_tot == 8 & ppl_in_home_allchild == 5 ~ 43470,
      ppl_in_home_tot == 8 & ppl_in_home_allchild == 6 ~ 42066,
      ppl_in_home_tot == 8 & ppl_in_home_allchild == 7 ~ 41709,
      ppl_in_home_tot >= 9 & ppl_in_home_allchild == 0 ~ 56621,
      ppl_in_home_tot >= 9 & ppl_in_home_allchild == 1 ~ 56895,
      ppl_in_home_tot >= 9 & ppl_in_home_allchild == 2 ~ 56139,
      ppl_in_home_tot >= 9 & ppl_in_home_allchild == 3 ~ 55503,
      ppl_in_home_tot >= 9 & ppl_in_home_allchild == 4 ~ 54460,
      ppl_in_home_tot >= 9 & ppl_in_home_allchild == 5 ~ 53025,
      ppl_in_home_tot >= 9 & ppl_in_home_allchild == 6 ~ 51727,
      ppl_in_home_tot >= 9 & ppl_in_home_allchild == 7 ~ 51406,
      ppl_in_home_tot >= 9 & ppl_in_home_allchild >= 8 ~ 49426
    ),
    inr_fpl = income_numeric / fpl_threshold
  ) %>% 
  ungroup()
```

## COPE data
```{r}
cope <-
  read_csv(cope_data) %>% 
  rename(current_num_child = current_num_bio_children, ID = cope_id) %>% 
  mutate(
    cohort = "COVID-19",
    past_mh_sa_treatment = dplyr::recode(
      past_mh_sa_treatment,
      "past treatment" = 1,
      "none" = 0
    )
  ) 
```

## Join BABIES and COPE
```{r}
d <-
  babies %>% 
  bind_rows(
    cope %>% 
      mutate(ID = as.character(ID))
  ) %>% 
  mutate(cohort = as.factor(cohort)) %>% 
  ungroup()
```

# Cleaning to enable comparisons

```{r}
# create variable for "White" and "Person of color"
d <-
  d %>% 
  mutate(
    poc = case_when(
      race_c == "White" ~ 0,
      mom_race == "White" ~ 0,
      !is.na(mom_race) & mom_race != "White" ~ 1,
      !is.na(race_c) & race_c != "White" ~ 1,
      mom_latinx == 1 ~ 1
    )
  ) 
```

```{r}
# recode job_status variable so BABIES and COPE bins are equivalent
d  <-
  d %>% 
  mutate(
    job_status = case_when(
      job_status == "Employed for wages" ~ "Employed",
      job_status == "fulltime paid job" ~ "Employed",
      job_status == "parttime paid job" ~ "Employed",
      job_status == "Self-employed" ~ "Employed",
      job_status == "Homemaker" ~ "Home caregiver",
      job_status == "home caregiver" ~ "Home caregiver",
      job_status == "Student" ~ "Student",
      job_status == "student" ~ "Student",
      job_status == "retired" ~ "Other",
      job_status == "other" ~ "Other",
      job_status == "Unable to work" ~ "Other",
      job_status == "Unemployed" ~ "Unemployed",
      job_status == "unemployed/looking" ~ "Unemployed"
    ),
    employed = if_else(
      !is.na(job_status) & job_status == "Employed", 1, 0
    )
  ) 
```

```{r}
# recode education variables so BABIES and COPE bins are equivalent
d <-
  d %>% 
  mutate(
    education = case_when(
      education == "7th-9th grade" ~ "< High school diploma",
      education == "10th-12th grade" ~ "< High school diploma",
      education == "Some high school, no diploma" ~ "< High school diploma",
      education == "High school/GED" ~ "High school/GED",
      education == "Partial college" ~ "Partial college, no degree",
      education == "Some college credit, no degree" ~ "Partial college, no degree",
      education == "Trade school/apprenticeship" ~ "Trade/technical/vocational training",
      education == "Trade/technical/vocational training" ~ "Trade/technical/vocational training",
      education == "2-year college degree" ~ "Associate degree",
      education == "Associate degree" ~ "Associate degree",
      education == "4-year college degree" ~ "Bachelor's degree",
      education == "Bachelor's degree" ~ "Bachelor's degree",
      education == "Graduate degree" ~ "Graduate degree",
    ),
    college_or_higher = if_else(
      !is.na(education) & education == "Bachelor's degree" | education == "Graduate degree",
      1, 0
    )
  ) 
```

```{r}
# create variable for partnered/married vs. not
d <-
  d %>% 
  mutate(
    partnered = case_when(
      marital_status_txt == "Married/domestic partnership" ~ 1,
      marriage_status == "Partner/married" ~ 1,
      TRUE ~ 0
    )
  ) 

```


```{r}
d <-
  d %>% 
  mutate(
    low_income = if_else(
      income_numeric <= 2 * fpl_threshold, 1, 0
    )
  )
```

# BABIES sample characteristics
```{r}
d %>% 
  filter(cohort == "Pre-pandemic") %>% 
  summarise_at(
    vars(
      mom_age,
      gestational_weeks,
      ppl_in_home_allchild,
      ppl_in_home_adults,
      inr_fpl,
      epds_total
    ),
    funs(mean, sd, min, max), na.rm = TRUE
  ) 
```

```{r}
d %>% 
  filter(cohort == "Pre-pandemic") %>% 
  count(first_pregnancy) %>% 
  mutate(per = n / sum(n))
```

```{r}
d %>% 
  filter(cohort == "Pre-pandemic") %>% 
  count(mom_race, mom_latinx) %>% 
  mutate(per = n / sum(n)) %>% 
  arrange(desc(n))

d %>% 
  filter(cohort == "Pre-pandemic") %>% 
  count(poc) %>% 
  mutate(per = n / sum(n)) %>% 
  arrange(desc(n))
```

```{r}
d %>% 
  filter(cohort == "Pre-pandemic") %>% 
  count(income_numeric > 100000) %>% 
  mutate(per = n / sum(n)) 

d %>% 
  filter(cohort == "Pre-pandemic") %>% 
  count(low_income) %>% 
  mutate(per = n / sum(n)) 
```

```{r}
d %>% 
  filter(cohort == "Pre-pandemic") %>% 
  count(education) %>% 
  mutate(per = n / sum(n)) %>% 
  arrange(desc(n))

d %>% 
  filter(cohort == "Pre-pandemic") %>%  
  count(college_or_higher) %>% 
  mutate(per = n / sum(n)) %>% 
  arrange(desc(n))
```

```{r}
d %>% 
  filter(cohort == "Pre-pandemic") %>% 
  count(job_status) %>% 
  mutate(per = n / sum(n)) %>% 
  arrange(desc(n))
```

```{r}
d %>% 
  filter(cohort == "Pre-pandemic") %>% 
  count(partnered) %>% 
  mutate(per = n / sum(n)) %>% 
  arrange(desc(n))
```

```{r}
d %>% 
  filter(cohort == "Pre-pandemic") %>% 
  count(past_mh_sa_treatment) %>% 
  mutate(per = n / sum(n))
```

# Distributions 
```{r}
d %>% 
  dplyr::select(
    cohort,
    `Age (years)` = mom_age,
    `Number of children` = ppl_in_home_allchild,
    `Gestational weeks` = gestational_weeks
  ) %>% 
  gather(
    key, 
    value,
    -cohort
  ) %>% 
  ggplot(aes(value, fill = fct_rev(cohort))) +
  geom_density(alpha = 1/2) +
  scale_fill_nejm() +
  theme_pubr() +
  theme(
    strip.text = element_text(size = 16),
    axis.title.y = element_text(size = 20),
    axis.title.x = element_text(size = 20),
    axis.text.y = element_text(size = 16),
    axis.text.x = element_text(size = 16),
    legend.title = element_text(size = 18),
    legend.text = element_text(size = 16),
    legend.position = "bottom"
  ) +
  labs(
    x = NULL,
    fill = "Cohort"
  ) +
  facet_wrap(.~key, scales = "free")


ggsave(
  "~/Box/lucy_king_files/BABIES/cope/figures/babies_cope_density_plots.png",
  dpi = 600,
  width = 10,
  height = 4
)
```

```{r}
counts_table <-
  d %>% 
  dplyr::select(
    ID,
    cohort,
    first_pregnancy,
    partnered,
    poc,
    low_income,
    college_or_higher,
    employed,
    past_mh_sa_treatment,
  ) %>% 
  gather(variable, value, -cohort, -ID) %>% 
  group_by(cohort, variable) %>% 
  mutate(
    sample_size = sum(!is.na(value))
  ) %>% 
  group_by(cohort, variable) %>% 
  summarise(
    sample_size = mean(sample_size),
    n = sum(value, na.rm = TRUE),
  ) %>% 
  mutate(
    Percent = n / sample_size,
    variable = recode_factor(
      variable,
      "college_or_higher" = "College or higher",
      "employed" = "Employed",
      "first_pregnancy" = "Primiparous",
      "low_income" = "Low income",
      "partnered" = "Married or partnered",
      "poc" = "Person of color",
      "past_mh_sa_treatment" = "Past mental health/substance treatment"
    )
  ) %>% 
  dplyr::select(-sample_size) %>% 
  ungroup()
```


# Test initial differences
```{r}
chisq.test(d$employed, d$cohort)
chisq.test(d$college_or_higher, d$cohort)
chisq.test(d$low_income, d$cohort) 
chisq.test(d$poc, d$cohort) 
chisq.test(d$past_mh_sa_treatment, d$cohort) 


chisq.test(d$first_pregnancy, d$cohort) # significant
chisq.test(d$partnered, d$cohort) # significant
d %>% 
  count(cohort, partnered) %>% 
  group_by(cohort) %>% 
  mutate(per = n / sum(n))


t.test(d$ppl_in_home_allchild ~ d$cohort)
t.test(d$mom_age ~ d$cohort) # significant
t.test(d$gestational_weeks ~ d$cohort) # significant
```

# Match samples

see: https://data.library.virginia.edu/getting-started-with-matching-methods/

```{r}
#first, only include COPE women in bay area counties
d %>% 
  count(ca_county)

d.1 <-
  d %>% 
  filter(
    !ca_county %in% c(
      "Monterey", 
      "Fresno", 
      "Orange", 
      "Humboldt", 
      "Sacramento", 
      "San Diego", 
      "Yolo"
    )
  ) %>% 
  mutate(
    exclude = if_else(cohort == "COVID-19" & is.na(ca_county), 1, 0)
  ) %>% 
  filter(exclude == 0)
```

## Optimal Nearest Neighbor

```{r}
# define what we want to match on; missing values on covariates not allowed

# missing values not allowed
d.2 <-
  d.1 %>% 
  filter(
    !is.na(ppl_in_home_allchild),
    !is.na(mom_age),
    !is.na(low_income),
    !is.na(college_or_higher),
    !is.na(past_mh_sa_treatment)
  )


# including education led to poorer match; therefore, not included
m.out <- matchit(
  cohort ~
    mom_age +
    gestational_weeks +
    ppl_in_home_allchild +
    first_pregnancy +
    partnered +
    employed +
    past_mh_sa_treatment,
  data = d.2,
  method = "optimal"
)
```

```{r}
# results of matching procedure
s.out <- summary(m.out, standardize = TRUE)
s.out
plot(
  s.out, 
  xlim = c(0, .8)
)
```

```{r}
# save matched data
d_matched <- match.data(m.out) 
```

## Coarsened exact matching

```{r}
m.out_cem <- matchit(
  cohort ~
    mom_age +
    gestational_weeks +
    ppl_in_home_allchild +
    first_pregnancy +
    partnered +
    employed +
    past_mh_sa_treatment,
  data = d.2,
  method = "cem"
)
```

```{r}
# results of matching procedure
s.out_cem <- summary(m.out_cem, standardize = TRUE)
s.out_cem

plot(
  s.out_cem, 
  xlim = c(0, .8)
)

```

```{r}
# save matched data
d_matched_cem <- match.data(m.out_cem) 
d_matched_cem %>% 
  count(cohort)
```
## Distributions of matched data 
```{r}
d_matched %>% 
  dplyr::select(
    cohort,
    `Age (years)` = mom_age,
    `Number of children` = ppl_in_home_allchild,
    `Gestational weeks` = gestational_weeks
  ) %>% 
  gather(
    key, 
    value,
    -cohort
  ) %>% 
  ggplot(aes(value, fill = fct_rev(cohort))) +
  geom_density(alpha = 1/2) +
  scale_fill_nejm() +
  theme_pubr() +
  theme(
    strip.text = element_text(size = 16),
    axis.title.y = element_text(size = 20),
    axis.title.x = element_text(size = 20),
    axis.text.y = element_text(size = 16),
    axis.text.x = element_text(size = 16),
    legend.title = element_text(size = 18),
    legend.text = element_text(size = 16),
    legend.position = "bottom"
  ) +
  labs(
    x = NULL,
    fill = "Cohort"
  ) +
  facet_wrap(.~key, scales = "free")


ggsave(
  "~/Box/lucy_king_files/BABIES/cope/figures/babies_cope_density_plots_matched.png",
  dpi = 600,
  width = 10,
  height = 4
)
```

```{r}
counts_table_matched <-
  d_matched %>% 
  dplyr::select(
    ID,
    cohort,
    first_pregnancy,
    partnered,
    poc,
    low_income,
    college_or_higher,
    employed,
    past_mh_sa_treatment,
  ) %>% 
  gather(variable, value, -cohort, -ID) %>% 
  group_by(cohort, variable) %>% 
  mutate(
    sample_size = sum(!is.na(value))
  ) %>% 
  group_by(cohort, variable) %>% 
  summarise(
    sample_size = mean(sample_size),
    n = sum(value, na.rm = TRUE),
  ) %>% 
  mutate(
    Percent = n / sample_size,
    variable = recode_factor(
      variable,
      "college_or_higher" = "College or higher",
      "employed" = "Employed",
      "first_pregnancy" = "Primiparous",
      "low_income" = "Low income",
      "partnered" = "Married or partnered",
      "poc" = "Person of color",
      "past_mh_sa_treatment" = "Past mental health/substance treatment"
    )
  ) %>% 
  dplyr::select(-sample_size) %>% 
  ungroup()

counts_table_matched
```
## Means, SDs, ranges of matched data
```{r}
d_matched %>% 
  group_by(cohort) %>% 
  summarise_at(
    vars(
      ppl_in_home_allchild,
      mom_age,
      gestational_weeks
    ),
    funs(mean, sd, min, max), na.rm = TRUE
  )
```

## Test differences after matching
```{r}
chisq.test(d_matched$employed, d_matched$cohort)
chisq.test(d_matched$college_or_higher, d_matched$cohort)
chisq.test(d_matched$low_income, d_matched$cohort) 
chisq.test(d_matched$poc, d_matched$cohort) 
chisq.test(d_matched$past_mh_sa_treatment, d_matched$cohort) 
chisq.test(d_matched$first_pregnancy, d_matched$cohort) 
chisq.test(d_matched$partnered, d_matched$cohort) 

t.test(d_matched$ppl_in_home_allchild ~ d_matched$cohort)
t.test(d_matched$mom_age ~ d_matched$cohort) 
t.test(d_matched$gestational_weeks ~ d_matched$cohort) 
```

# Test Aim 1

H1: The COPE sample of pregnant mothers will have higher mean EPDS than the BABIES/pre-COVID pregnant mothers.

# Center factors
```{r}
d_matched <-
  d_matched %>% 
  mutate_at(
    vars(college_or_higher, employed, poc, first_pregnancy, past_mh_sa_treatment, partnered, low_income),
    as.factor
  )

contrasts(d_matched$cohort)

contrasts(d_matched$low_income) <- c(-.5, .5)

contrasts(d_matched$college_or_higher) <- c(-.5, .5)

contrasts(d_matched$employed) <- c(-.5, .5)

contrasts(d_matched$first_pregnancy) <- c(-.5, .5)

contrasts(d_matched$past_mh_sa_treatment) <- c(-.5, .5)

contrasts(d_matched$partnered) <- c(-.5, .5)
```

# Unadjusted t-test
```{r}
t.test(d_matched$epds_total ~ d_matched$cohort)
cohens_d(d_matched$epds_total, d_matched$cohort)
```

# Adjusted OLS
```{r}
lm_a1.1 <- 
  lm(
  epds_total ~
    cohort +
    low_income +
    college_or_higher +
    employed +
    first_pregnancy +
    past_mh_sa_treatment +
    partnered +
    poc +
    scale(mom_age, scale = FALSE) +
    scale(gestational_weeks, scale = FALSE) +
    scale(ppl_in_home_allchild, scale = FALSE),
  data = d_matched
)
summary(lm_a1.1)
model_parameters(lm_a1.1, standardize = "refit")
```

# Unadjusted chi-square
```{r}
d_matched <-
  d_matched %>% 
  mutate(
    epds_above_cutoff = if_else(
      epds_total >= 11, 1, 0
    )
  )

chisq.test(d_matched$epds_above_cutoff, d_matched$cohort)

epds_cutoff_table <- 
  d_matched %>% 
  count(cohort, epds_above_cutoff) %>% 
  group_by(cohort) %>% 
  mutate(per = n / sum(n))

epds_cutoff_table

# relative risk of COPE vs. BABIES
depression_cope <- epds_cutoff_table[2, 4] %>% pull()
depression_babies <- epds_cutoff_table[4, 4] %>% pull()
RR <- depression_cope / depression_babies
RR
```

## Visualize main effect
```{r}
d_matched %>% 
  ggplot(aes(x = fct_rev(cohort), y = epds_total, fill = cohort)) +
  geom_flat_violin(
    aes(fill = cohort),
    position = position_nudge(x = .21, y = 0), 
    adjust = 1.5, 
    trim = TRUE, 
    alpha = .65, 
    colour = NA
  ) +
  geom_point(
    aes(x = as.numeric(fct_rev(cohort))-.26, y = epds_total, colour = cohort),
    position = position_jitter(width = .05), 
    size = 2,
    alpha = .65,
    shape = 20
  ) +
  geom_boxplot(
    aes(x = fct_rev(cohort), y = epds_total, fill = cohort),
    outlier.shape = NA, 
    alpha = .65, 
    width = .4, 
    colour = "black"
  ) +
  # remove grid lines, etc.
  theme_pubr() +
  # adjust size and position of labels
  theme(
    axis.title.y = element_text(size = 20),
    axis.title.x = element_text(size = 20),
    axis.text.y = element_text(size = 16),
    axis.text.x = element_text(size = 16),
    # remove legend
    legend.position = "none"
  ) +
  # change names of labels
  labs(
    x = "Cohort",
    y = "Prenatal depressive symptoms\n(EPDS)"
  ) +
  # add color pallete
  scale_color_nejm() +
  scale_fill_nejm() 

ggsave(
  "~/Box/lucy_king_files/BABIES/cope/figures/babies_cope_epds.jpeg",
  width = 6,
  height = 5,
  dpi = 1000
)
```
# Compare participants in COVID-19 subset to participants in full cohort
```{r}
d_matched <-
  d_matched %>% 
  mutate(
    matched_subset = if_else(
      cohort == "COVID-19", 1, 0
    )
  )

cope <-
  cope %>% 
  left_join(
    d_matched %>% 
      filter(cohort == "COVID-19") %>% 
      dplyr::select(ID, matched_subset) %>% 
      mutate(ID = as.numeric(ID)),
    by = "ID"
  )

cope <-
  cope %>% 
  mutate(
    matched_subset = as.factor(
      if_else(is.na(matched_subset), 0, matched_subset)
    )
  ) 
```

```{r}
cope %>% 
  group_by(matched_subset) %>% 
  summarise_at(
    vars(epds_total),
    funs(mean, sd), na.rm = TRUE
  )

t.test(cope$epds_total ~ cope$matched_subset)
cohens_d(cope$epds_total, cope$matched_subset)
```
# EPDS ~ gestational age
```{r}
d %>% 
  ggplot(aes(gestational_weeks, epds_total, color = fct_rev(cohort))) +
  geom_point(alpha = 1/2, size = 1.5) +
  geom_smooth(method = "lm", size = 2.5) +
  scale_color_aaas() +
  theme_pubr() +
  theme(
    strip.text = element_text(size = 18),
    axis.title.y = element_text(size = 20),
    axis.title.x = element_text(size = 20),
    axis.text.y = element_text(size = 16),
    axis.text.x = element_text(size = 16),
    legend.text = element_text(size = 20)
  ) +
  labs(
    color = NULL,
    x = "Gestational weeks in pregnancy",
    y = "Prenatal depressive symptoms\n(EPDS)"
  ) 

ggsave(
  "~/Box/lucy_king_files/BABIES/cope/figures/gestation_weeks_epds_cohort.png",
  width = 7,
  height = 5,
  dpi = 1000
)
```

# Export data to share
```{r}
matched_COPE_IDs <-
  cope %>% 
  dplyr::select(ID, matched_subset)

write_csv(
  matched_COPE_IDs,
  "~/Box/lucy_king_files/BABIES/cope/matched_COPE_IDs_n82_20201207.csv"
)

matched_IDs <-
  d_matched %>% 
  mutate(matched_subset = "matched_subset") %>% 
  dplyr::select(ID, matched_subset)

d_share <-
  d %>% 
  dplyr::select(
    ID,
    cohort,
    epds_total,
    mom_age,
    mom_race,
    mom_latinx,
    job_status,
    employed,
    education,
    college_or_higher,
    income_numeric,
    low_income,
    inr_fpl,
    poc,
    past_mh_sa_treatment,
    first_pregnancy,
    partnered,
    ppl_in_home_allchild,
    ppl_in_home_adults,
    gestational_weeks,
    ca_county
  ) %>% 
  left_join(
    matched_IDs, by = "ID"
  )

var_label(d_share) <-
  list(
    cohort = "Women pregnant prior to or during the COVID-19 pandemic",
    epds_total = "Total score on Edinburgh Postnatal Depression Scale",
    mom_age = "Participant's age at assessment",
    mom_race = "Participant's self-reported race",
    mom_latinx = "Did participant self-reported Hispanic/Latinx ethnicity?",
    job_status = "Is participant employed for wages?",
    education = "Participant's self-reported level of education",
    college_or_higher = "Did participant complete at least a 4-year college degree?",
    income_numeric = "Participant's self reported income (originally reported in bins; here, midpoint of bin to allow calculation of income-to-needs ratio)",
    inr_fpl = "Income-to-needs ratio based on federal poverty line",
    poc = "Did participant identify as a person of color?",
    past_mh_sa_treatment = "Did participant receive treatment for mental health difficulties or substance abuse in the past?",
    first_pregnancy = "Is this the participant's first pregnancy?",
    partnered = "Does the participant have a partner/significant other?",
    ppl_in_home_allchild = "Total number of children in participant's household",
    ppl_in_home_adults = "Total number of adults in participan'ts household",
    ca_county = "California county in which participant resides (missing for pre-pandemic participants because all were from San Francisco Bay Area counties",
    matched_subset = "Is the COVID-19 participant part of the subset matched to the pre-prandemic participants?",
    gestational_weeks = "Number of gestational weeks into pregnancy (calculated based on due date)"
  )

metadata(d_share)$name <- "Analysed data for Aim 1 of \"Pregnancy during the pandemic\\: The impact of COVID-19-related stress on risk for prenatal depression\" (King et al. 2020)"

write_csv(
  d_share,
  "~/Desktop/BABIES/COPE/cope_git/analyzed_data_Aim1.csv"
)

```


